\chapter{Conclusions}
\label{ch:conclusions}
In this work I developed an electrical motor handling hardware/software
solution with modularity and extensibility in mind. In this very final chapter
I demonstrate how ammc can be used in practice and I explain how it could be
developed further in the future.

\section{Example setup}

\subsection{Hardware}
The hardware required for the example setup given here is the following:
\begin{itemize}
  \item One \emph{ATMega2560} microcontroller unit for the master controller
  \item One \emph{ATMega2560} microcontroller unit for the slave controller
  \item Two $4.7 k\Omega$ resistors for the I2C open-drain
  \item A $100 pF$ capacitor to inhibit the reset-over-serial capabilities of
    the ATMega2560 board
  \item A 150 rpm dc motor with embedded two-phase encoder
  \item An L293 h-bridge motor driver
  \item A power supply compatible with the h-bridge and motor voltage and
    power requirements (e.g.\ $12 V,\ 24 W$)
\end{itemize}

The hardware setup must be performed as shown in figures \ref{img:master-sch}
and \ref{img:slave-sch}.
The L293 driver must be powered with two independent power supplies, one at
$12V$ (for the motor output amplifier) and one at $5V$ (for the driver logic
circuit. The AVR microcontrollers can be powered using their USB sockets. The
I2C reference voltage can be set by using a $5 V$ output pin of the master
controller board.

\subsection{Software}
The software can be built and eventually flashed using GNU Make. the
\emph{avr-gcc} toolchain and the \emph{avr-libc} library are required for the
build, while \emph{avrdude} is required for flashing.

The \texttt{make} command will build all the ammc software, i.e.\ the master
and slave controllers firmwares and the client application. The \texttt{make
master-flash} and the \texttt{make slave-flash} commands will flash the master
and slave firmwares into their respective microcontrollers. The compiled client
application will be located under the \texttt{client/target} directory.

\section{Future work}

\subsection{Extended connectivity}
The extension of the master controller connectivity would make ammc suitable
for whole new range of application possibilities. For example, \emph{LoRaWAN}
and similar communication vectors would make ammc a motor management system
suitable for dc motor distributed in a wide-area. The use of multiple wired
low-level communication means such as SPI or raw serial ports would make ammc
suitable for a multi-master, multi-group setup.

\subsection{Programming API}
The client application offers a non-interactive mode for basic scriping
functionality. Thus it is good for basic automation, it is quite limited when
more complex features are required, such as an advanced use of logic or control
flow capabilities. For this purpose, the implementation of a programming API,
eventually with bindings in various languages, would make the ammc
functionality directly available to the programmer as high-level functions.
